---
title: "A minimal notebook to analyse test data\n"
author: "Denis MOTTET"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: false
    number_sections: true
---

# Objective
The goal of this notebook is to illustrate a minimal data analysis pipeline, from a data file to a result chart.   
We shall count the number of individuals and save a graph of their gender distribution. 


NOTE: this notebook follows the logic of the `.gitignore` file to ensure the confidentiality of personal data from human volunteers while allowing the testing of methods on pseudonymized data that we have made public.  
- the `data` folder is not shared by Git, except for files in the `data/test_data/` subfolder.  
- the `results` folder is not shared by Git, except for files in the `results/test_results/` subfolder.  


# Load the data
The data file is provided in the `data/test_data/` folder.  
The csv file contains one individual per line and several columns, with the column name appearing in the first line of the file.

```{r load_data, echo=TRUE, message=FALSE, warning=FALSE}

library(effsize)
library(readxl)
# Load the data from the CSV file
data <- read_xlsx("data/Global_Jump_Height.xlsx")
# Display the first few rows of the dataset
head(data)
```
```{r}
# Calcul du Delta
data$Delta <- data$CMJ_Mean - data$SJ_Mean

mean(data$Delta, na.rm = TRUE)
sd(data$Delta, na.rm = TRUE)
```

```{r}
# Verification de la normalité
shapiro_result <- shapiro.test(data$Delta)
shapiro_result
```

# Statistical Analysis of the data (paired t-test)
```{r}
# Test t apparié
t.test(data$CMJ_Mean, data$SJ_Mean, paired = TRUE)
```


# Taille d'effet
```{r}
effect_size <- cohen.d(data$CMJ_Mean, data$SJ_Mean, paired = TRUE)
effect_size
```


