---
title: "Projet Analyse du Mouvement\n"
author: "Chahine ABDEL-HAMID"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
    highlight: tango
    number_sections: true
---

# Statistical analysis in R

This notebook presents the **statistical analysis** of the movement analysis project.

The dataset used here was previously created in Python and saved as:

- `data/Global_Jump_Height.xlsx`

The analysis focuses on the variable:

- **Jump Height (Flight Time) [cm]**

for the two jump conditions:

- **Squat Jump (SJ)**
- **Countermovement Jump (CMJ)**

The purpose of this R notebook is to:

- describe the data,
- compute the paired difference between CMJ and SJ,
- assess the normality of the paired differences,
- perform the appropriate paired statistical test,
- estimate the effect size.

# Load required packages and Import the cleaned dataset 

The cleaned file generated in Python is imported below.

```{r,warning=FALSE}

library(readxl)
library(effsize)
library(dplyr)
library(knitr)


df <- read_xlsx("data/Global_Jump_Height.xlsx")
```


# Convert variables to numeric and Compute the paired difference 

The two main variables (SJ_Mean and CMJ_Mean) must be treated as numeric variables.

Then, a new variable called Delta is created:

$Delta = CMJ_mean - SJ_Mean$

This variable represents the individual gain in jump height between the two conditions.
It is central to the statistical reasoning because the paired comparison relies on the within-subject difference.

```{r,warning=FALSE}

df$SJ_Mean <- as.numeric(df$SJ_Mean)
df$CMJ_Mean <- as.numeric(df$CMJ_Mean)


df$Delta <- df$CMJ_Mean - df$SJ_Mean
```

# Table of Individual values 

Before summarizing the results at the group level, it is useful to present the individual values.

This table shows, for each participant:

- the mean jump height in SJ,
- the mean jump height in CMJ,
- the paired difference (CMJ - SJ).

```{r,warning=FALSE}
individual_table <- df %>%
  select(Patient, SJ_Mean, CMJ_Mean, Delta)

kable(
  individual_table,
  digits = 2,
  caption = "Individual jump height values and paired differences"
)
```

A **positive value** in the Delta column indicates that the participant **jumped higher in the CMJ** condition than in the SJ condition.
A **negative value** would indicate **the opposite**.

#  Descriptive statistics 

The next step is to summarize the data at the group level.

We compute:

- the mean and standard deviation for SJ_Mean,
- the mean and standard deviation for CMJ_Mean,
- the mean and standard deviation for Delta.

These descriptive statistics provide an overview of the performance in each condition.

```{r,warning=FALSE}
mean_SJ <- mean(df$SJ_Mean, na.rm = TRUE)
sd_SJ   <- sd(df$SJ_Mean, na.rm = TRUE)

mean_CMJ <- mean(df$CMJ_Mean, na.rm = TRUE)
sd_CMJ   <- sd(df$CMJ_Mean, na.rm = TRUE)

mean_delta <- mean(df$Delta, na.rm = TRUE)
sd_delta   <- sd(df$Delta, na.rm = TRUE)
```

The descriptive statistics are displayed below in a clean summary table.

```{r,warning=FALSE}
desc_table <- data.frame(
  Condition = c("SJ", "CMJ", "CMJ - SJ"),
  Mean = c(mean_SJ, mean_CMJ, mean_delta),
  SD = c(sd_SJ, sd_CMJ, sd_delta)
)

kable(
  desc_table,
  digits = 2,
  caption = "Descriptive statistics for jump height"
)

```

This table allows a first interpretation of the results.

If the mean value is higher in CMJ than in SJ, this already suggests a performance advantage in the countermovement condition.

However, descriptive statistics alone are not sufficient to conclude whether the observed difference is statistically reliable.

# Normality of paired differences 

Because the main inferential test planned in this project is a paired Student t-test, we first need to verify whether the paired differences are approximately normally distributed.

Importantly, the normality assumption is tested on:

$Delta = CMJ - SJ$

and not separately on *SJ_Mean* and *CMJ_Mean*.

This is because the paired t-test is based on the distribution of the within-subject differences.

```{r,warning=FALSE}

shapiro_result <- shapiro.test(df$Delta)
shapiro_result
```

**Interpretation of the normality test**

The null hypothesis of the Shapiro-Wilk test is that the data are *normally distributed*.

If p > 0.05, the normality assumption is considered acceptable.
If p < 0.05, normality would be questioned.

In this project, the p-value is greater than 0.05, which means that the normality assumption is not violated.
This supports the use of a paired t-test for the next step.

# Paired t-test 

```{r,warning=FALSE}

ttest_result <- t.test(df$CMJ_Mean, df$SJ_Mean, paired = TRUE)
ttest_result
```

**Interpretation of the paired t-test**

The paired t-test provides several important pieces of information:

- the t statistic,
- the degrees of freedom,
- the p-value,
- the confidence interval of the mean difference,
- the estimated mean difference between conditions.

If the p-value is below 0.05, the difference between CMJ and SJ is considered statistically significant.

Here, the result indicates that *jump height is significantly greater in the CMJ condition than in the SJ condition*.


#  Effect size

A statistically significant p-value tells us that a difference exists, but it does not tell us how large that difference is in practical terms.

To address this, we compute Cohen's d, which provides a standardized effect size.
```{r,warning=FALSE}

effect_size <- cohen.d(df$CMJ_Mean, df$SJ_Mean, paired = TRUE)
effect_size
```
**Interpretation of the effect size**

Cohen's d is commonly interpreted as follows:

- around 0.2: small effect
- around 0.5: medium effect
- around 0.8 or more: large effect

In this project, the *effect size is large*, which indicates that the *difference between SJ and CMJ is not only statistically significant, but also substantial in magnitude*.


# Full Statistical analysis 

This table provides a compact summary of the main statistical conclusions:

- the normality assumption was respected,
- the paired difference was statistically significant,
- the effect size was large.

```{r,warning=FALSE}
infer_table <- data.frame(
  Test = c("Shapiro-Wilk", "Paired t-test", "Effect size"),
  Statistic = c(
    round(unname(shapiro_result$statistic), 4),
    round(unname(ttest_result$statistic), 4),
    round(effect_size$estimate, 4)
  ),
  `p-value` = c(
    round(shapiro_result$p.value, 4),
    signif(ttest_result$p.value, 3),
    NA
  ),
  Interpretation = c(
    "Normality assumption respected",
    "Significant difference between CMJ and SJ",
    "Large effect"
  )
)

kable(
  infer_table,
  caption = "Inferential statistics"
)
```

The statistical analysis shows that:

- mean jump height was *34.70 ± 2.98 cm* in the SJ condition,
- mean jump height was *38.27 ± 3.26 cm* in the CMJ condition,
- mean paired difference (CMJ - SJ) was *3.57 ± 2.02 cm*.

The paired differences were normally distributed:

Shapiro-Wilk: W = 0.9762, p = 0.9638

The paired t-test showed a statistically significant difference between conditions:

- t(11) = 6.1163
- p = 7.56e-05
- 95% CI = [2.28 ; 4.85]

The effect size was large:

- Cohen's d = 1.13

Taken together, these results indicate that **jump height was significantly greater in the CMJ condition than in the SJ condition**.

Because the CMJ includes a preparatory eccentric phase whereas the SJ starts from a static position, this difference can be interpreted as an indirect functional indicator of *stretch-shortening cycle contribution*.

In this sample of athletes with patellar tendinopathy, the expected performance advantage of the CMJ therefore remained present at the group level.